---
title: "Visualization Project"
author: "Arka Roy"
date: "`2023-09-24`"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```
## Data Set- Mercedes Used Car Listing
### Introduction
  This is a data set on used mercedes cars models depicting the price,mileage and various other aspects of these      cars.Various factors determine cost of an used car given a certain model like how many miles it has run, and what   year the car was manufactured. Additionally, there are various cheap and expensive models for each brand,           especially for a Mercedes. Thus, we will be analyzing these data to see what the prices would be.

### Data Description

* The data set contains 9 columns and 13119 rows.
* It has no missing data.
* Column 'model' tells the car model.
* Column 'Year' gives the year of manufacturing of the models which ranges from 1970 to 2020 with the majority        models being manufactured in 2018.
* Column 'price' gives the selling prices of these used cars which ranges from 650 pounds to 159999 pounds with the majority models being sold at 24699 pounds.
* Column 'Transmission' tells us the type of transmission like automatic,manual or semi-transmission.
* Column 'mileage' tells us the no of miles the car has run before being sold. It ranges from as low as 1 mile to as high  as 259000 miles with the most frequent being 15189 miles.
* Column 'Fueltype' have three data types- 'Petrol', 'Diesel','Hybrid'.
* Column 'tax' tells the government taxation rate applicable on these cars when they were sold.
* Column 'mpg' tells us the mileage per gallon for these models when they were sold.
* Column 'engineSize' gives information about volume of the engine.

```{r,echo=FALSE}
library(readr)
mydata=read.csv("/Users/arkaroy/Desktop/Kaggle/merc.csv")
head(mydata,n=5)
```

### Exploratory Data Analysis

 
```{r,echo=FALSE}
count1=table(mydata$transmission)
print(count1)

```


* Initial Exploration shows that among the total models sold, Highest were of Semi-Auto Transmission followed by Automatic and Manual.

* If we explore the registration year with re-sale pricing for these transmission types, we can easily observe that apart from certain out liers, mostly semi-Auto type transmission with latest registration year is more concentrated at higher re-sale price zone. The price for automatic type is more distributed. One can also infer that semi-automatic transmission type is a recent product from Mercedes since the majority registrations were only 2 decades old.


```{r, echo=FALSE}
library(ggplot2)
ggplot(mydata)+geom_point(aes(x = year, y = price,colour=transmission),size=0.2)+ggtitle("Variation of re-sale price of all models based on transmission type")+coord_trans(x="log",y="log")+facet_wrap(~transmission,ncol=3)+ theme(axis.text.x = element_text(angle = 45),axis.text.y = element_text(angle = 0))


```


* Similarly upon observation,we can infer that petrol variants have re-sale values more distributed with certain high re-sale price models than diesel.



```{r, echo=FALSE}
library(ggplot2)
  ggplot(mydata)+geom_point(aes(x = year, y = price,colour=fuelType),size=0.5)+ggtitle("Variation of re-sale price of models with respect to fuel type")+
  coord_trans(x="log",y="log")+facet_wrap(~fuelType,ncol=5) + theme(axis.text.x = element_text(angle = 90))


```

* Now we will try to explore and find out the reason behind the high price of certain petrol variant models.
  we will look into a filtered data set with only petrol variants as fueltype.
* So we will first filter the fuelType to "Petrol" variants only.
  
  
```{r, echo=FALSE}
mydata_petrol <- dplyr::filter(mydata, fuelType=="Petrol")
head(mydata_petrol)
```

* Now we have a filtered data set where we will be exploring the reason behind high prices of certain petrol variants.


* We will try to explore whether this high re-sale price of petrol variants is connected to the transmission type.

```{r, echo=FALSE}
ggplot(mydata_petrol)+geom_point(aes(x = year, y = price,colour=transmission),size=0.5)+ggtitle("Price Variation of only petrol variant models based on transmission type")   + coord_trans(x="log",y="log")   +theme(axis.text.x = element_text(angle = 45))+facet_wrap(~transmission,ncol=3)
```

* Clearly, for petrol as fuel type, transmission type is not the factor here as both automatic and semi-auto are falling in the same price range.

* Let's explore the model types to find out whether it is dependent on the models.
  first we will see what are the different model types and how their sale depended on fuel type.
  
  
```{r,echo=FALSE}

ggplot(data = mydata, 
       aes(x = model, fill = fuelType)) + 
  geom_bar() +  ggtitle("Model Re-sale count with respect to fuel type")+
   theme(axis.text.x = element_text(angle = 90))

```


* So there are total 27 different models of different fuel types. Among all of them C class model has been re-sold   the highest which is a compact sedan within an affordable price range followed by A-Class which is an hatch        back.
* Clearly C-Class model with diesel as fuel type surpasses the sale count when compared to its immediate             competitor-petrol fuel type.
* Also one can observe that in almost every model type, diesel variety re-sale count exceeds that of petrol. 

* Now lets find out how prices of different model types varies in petrol segment. We will try to analyze the mean    price of the different models to get an idea.

* First let calculate the mean price for each model type.


```{r,echo=FALSE}
mean_prices <- aggregate(mydata_petrol$price, by = list(mydata_petrol$model), FUN = mean)
colnames(mean_prices) <- c("Model", "Mean_Price")
head(mean_prices)
```

* Now let's find out which model types among petrol variant has higher prices.

```{r,echo=FALSE}
library(ggplot2)
ggplot(mean_prices)+geom_point(aes(x = Model, y = Mean_Price),color="red") +
  labs(x = "Car Model", y = "Mean Price") +
  ggtitle("Mean Prices of Used Mercedes Car Models of petrol variants") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```

* So clearly the mean re-sale price of G-class with petrol variant is very high when compared to other models,    which is because it is an SUV wagon from the house of Mercedes. The lowest price point is that of CLK followed by CLC class and 230.
  
 * Lets find out what is the scenario for models with diesel as fuel type.
 
```{r,echo=FALSE}
mydata_diesel <- dplyr::filter(mydata, fuelType=="Diesel")
mean_pricesDiesel <- aggregate(mydata_diesel$price, by = list(mydata_diesel$model), FUN = mean)
colnames(mean_pricesDiesel) <- c("Model", "Mean_Price")
library(ggplot2)
ggplot(mean_pricesDiesel)+geom_point(aes(x = Model, y = Mean_Price),color="blue") +
  labs(x = "Car Model", y = "Mean Price") +
  ggtitle("Mean Prices of Used Mercedes Car Models of diesel variants") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

```


* Here also we can see that price point of G-CLASS dominates the other models which was expected. But the price of 
  Diesel variant of G class is lower significantly than that of the petrol variant which answers our question as to   which petrol models have higher price point. So the reason is solely dependent on the model type.
  
* We can also see that other models like A CLASS and C class of Diesel variants also have lower price points when    compared to Petrol variants.
.
  
* Now we will try to understand how price of C Class which is the most re-sold car varies with different factors

```{r,echo=FALSE}
Cclass_data <- subset(mydata, model == " C Class")

library(ggplot2)
ggplot(Cclass_data)+geom_point(aes(x = year, y = price,colour=transmission),size=0.5)+ggtitle(" C class price vs transmission for different fuel type") +coord_trans(x="log",y="log")+facet_wrap(~fuelType,ncol=5)

``` 


* Clearly , we can observe that in case of C-class , the manual mode of transmission has slowly phased out   from the re-sale market and is being replaced by the Semi-auto type transmission

* The re-sale price point for both Diesel and Petrol variants of C class slowly has increased manifolds over last    15 years.

* The semi-automatic transmission type is slowly dominating the automatic type in the re-sale market both in Petrol   and Diesel,whereas for hybrid fuel-type,it is still dominated by automatic type.

* Another important observation for both A-class and C -class is that both entered the re-sale market well after     2005.

 * Now lets see how the prices of C class varied with respect to mileage
 
```{r,echo=FALSE}
library(ggplot2)
ggplot(Cclass_data)+geom_point(aes(x = mileage, y = price,colour=transmission),size=0.5)+ggtitle(" C class mileage vs  transmission for different fuel type") +facet_wrap(~fuelType,ncol=5)+theme(axis.text.x = element_text(angle = 90, hjust = 1))
```


* As expected we can see that re-sale price point dropped drastically when the mileage increased.

* The striking outcome is that in the re-sale data, majority petrol variant C -class models have a mileage in the range of 0 to 50000 mileas whereas that of Diesel is 25000 to 75000 miles.

* So the reason behind higher price points of petrol variants over diesel variants is clearly because overall        mileage of the petrol variant models is lower than that of C-class models.


### Result


1. The Mercedes resale car market really flourished after 2005 where although initially the automatic transmission    type dominated alongwith manual, they are slowly being overtaken by semi-auto type transmission.

2. A class and C class are the most resold models of Mercedes among which C-Class is the highest.

3. For every model except SL Class, Diesel fueltype variants have dominated the re-sale market over Petrol type.

4. G-Class models have the highest price. In general, for every model, petrol variants in the re-sale market has 
   a higher mean price than diesel variants.
   
5. Manual mode of transmission has slowly phased out from the re-sale market.

6. Re-sale prices of A-Class and C-Class have increased manifolds in the recent year and generally price points are    higher when the mileage is lower and registration year is not old.












### Conclusion


Mercedes re-sale car market in the Europe has really flourished over the years since 2005 attracting young executives and working professionals to get hold of the prestigious Mercedes Benz cars. Most preferred model types are A class and C class which are kind of entry level models for Mercedes with an affordable price of around 50,000 pounds.Although the electric era has boomed, for Mercedes, the preferred fuel type is still Diesel followed by Petrol and the preferred transmission has become semi-automatic. Manual transmission has lost its charm in the market and is slowly getting phased out. Inspite of the advent of covid in 2020, the market was steadfast and in fact prices increased during this period. 
